---
title: "**First Bayes: Beta-Binomial**"
author: ""
output: 
  flexdashboard::flex_dashboard:
    theme: lumen
    vertical_layout: fill
    orientation: columns
    source_code: https://github.com/GiankDiluvi/first-bayes-r
runtime: shiny
---

<style type="text/css">

.sidebar {  /* sidebar  */
   font-size: 14px;
   font-family: Helvetica;
   }
.navbar-inverse .navbar{
  font-family: Helvetica;
}
.navbar-inverse .navbar-brand{
  font-family: Helvetica;
  color: #000000;
  pointer-events: none;
}
</style>




```{r global, include=FALSE, warning = FALSE}
# preamble ###
# basics
library(flexdashboard)
library(tidyverse)
# plotting
library(ggplot2)
ggplot2::theme_set(theme_classic())
library(viridis)
library(patchwork)
# markdown
library(knitr)
library(kableExtra)
# shiny + html
library(shiny)
withMathJax()
library(DT)
# extras
#library(pscl)
set.seed(2020)
```


```{r}
# taken from package pscl (not installed in shiny-servetr.stat.ubc)
# code: https://github.com/cran/pscl/blob/master/R/betaHPD.r
# author: Simon Jackman
# todo: install package pscl in stat server and use that instead
betaHPD <- function(alpha,beta,p=.95,plot=FALSE,xlim=NULL,debug=FALSE){
  
  if(is.na(p) | is.nan(p) | p > 1 | p < 0)
    stop("p not between 0 and 1\n")
  
    #if(alpha<=1 | beta <=1)
    #  stop("betaHPD only implemented for alpha and beta both > 1\n")

  ## initialize internal logical flags
  compute <- TRUE
  swap <- FALSE

  if(alpha==beta){
    if(debug)
      cat("symmetric case, alpha=",alpha,"beta=",beta,"\n")
    out <- qbeta((1 + c(-1,1)*p)/2,
                 alpha,beta)
    compute <- FALSE
  }
  if(alpha>beta){
    swap <- TRUE
    alphaStar <- beta
    betaStar <- alpha
  }
  else if(beta>alpha){
    swap <- FALSE
    alphaStar <- alpha
    betaStar <- beta
  }
  if(debug)
    cat("swap=",swap,"\n")
  
  func <- function(x0,alpha,beta){
    y0 <- dbeta(x0,alpha,beta)
    p0 <- pbeta(x0,alpha,beta)
    x1 <- qbeta(p0+p,alpha,beta)
    y1 <- dbeta(x1,alpha,beta)
    out <- abs(y0-y1)
    out
  }
  
  if(compute){
    foo <- try(optimize(f=func,alpha=alphaStar,beta=betaStar,
                        tol=.Machine$double.eps^(.6),
                        interval=c(.Machine$double.eps,
                          qbeta(1-p,
                                alphaStar,betaStar))))
    if(inherits(foo,"try-error")){
      warning("optimization in betaHPD failed\n")
      out <- rep(NA,2)
    }
    else{
      if(debug){
        cat("results of optimization:\n")
        print(foo)
      }
      out <- c(foo$minimum,
               qbeta(pbeta(foo$minimum,alphaStar,betaStar)+p,
                     alphaStar,betaStar)
               )
    }
    if(swap){
      out <- 1-out
      out <- sort(out)
      if(debug){
        cat("swapped back\n")
        print(out)
      }
    }
  }

  ## plotting
  if(plot & all(!is.na(out))){
    xseq <- NULL
    if(length(xlim)==2 & all(!is.na(xlim))){
      if(xlim[2]>xlim[1] & xlim[1] >= 0 & xlim[2] <= 1){
        xseq <- seq(xlim[1]+(.Machine$double.eps^(.25)),
                    xlim[2]+(.Machine$double.eps^(.25)),
                    length=1000)
      }
    }
    if(is.null(xseq))
      xseq <- seq(min(qbeta(.0001,alpha,beta),out[1]),
                  max(qbeta(.9999,alpha,beta),out[2]),
                  length=1000)
    
    plot(xseq,dbeta(xseq,alpha,beta),
         xlab=expression(theta),
         ylab="",
         axes=F,
         type="n")
    axis(1)

    ## get polygon for HDR
    dseq <- seq(out[1],out[2],length=250)
    fx <- dbeta(dseq,alpha,beta)
    polygon(x=c(out[1],dseq,rev(dseq)),
            y=c(0,fx,rep(0,250)),
            border=F,col=gray(.45))
    lines(xseq,dbeta(xseq,alpha,beta))
  }

  out
}
```



```{r datasets}
# create data sets for analyses ###
dog <- tibble(x = c(rep(1, 33), rep(0, 3)))        # dog identifies 33 of 36 cancer tests; https://gut.bmj.com/content/60/6/814
delta <- tibble(x  = c(0,1,1,1,1,1,1,1,1,1,1,1,1)) # delta IV heavy launches
```



**Prior distribution**
=====================================



Column {.sidebar data-width=350}
-------------------------------------

The prior distribution of $p$ is $\mathrm{Beta}(a, b)$, and it characterizes your prior knowledge of $p$. The values $a$ and $b$ are called the *hyperparameters*. 


**Select hyperparameters:**


```{r binom.hyperparams}

sliderInput("alpha", label = "a",
            min = 0, max = 10, value = 4, step = 0.1)

sliderInput("beta", label = "b",
            min = 0, max = 10, value = 4, step = 0.1)


```

<hr>


With the values of $a$ and $b$ that you chose, the prior expected value $\mathbb{E}[p]$ and variance $\mathrm{Var}[p]$ are shown below.


```{r prior.mean.var}

renderTable({
  
  # read inputs
  alpha <- input$alpha
  beta <- input$beta
  
  # calculate mean and variance
  mean <- alpha / (alpha + beta)
  var <- alpha * beta / ((alpha + beta)^2 * (alpha + beta + 1))
  
  # create table
  summary <- tibble(Mean = round(mean, 2),
                    Variance = round(var, 4))
  
  print(summary)
  
})

```





Column
-------------------------------------


### 
<!--
The plot below displays the density of the prior distribution with the hyperparameters that you chose.
-->

```{r prior}

renderPlot({
  
  # read inputs
  alpha <- input$alpha
  beta <- input$beta
  dat.name <- input$binom.data
  phat <- input$phat
  n <- input$n
  
  
  
  
  # calculate mean and variance
  mean.prior <- alpha / (alpha + beta)
  var.prior <- alpha * beta / ((alpha + beta)^2 * (alpha + beta + 1))
  
  
  
  ggplot(tibble(x = c(0, 1)), aes(x)) +
    stat_function(fun = dbeta, 
                  args = list(shape1 = alpha, shape2 = beta),
                  color = "#482173FF",
                  size = 2,
                  n = 1000) + #adds prior density plot
    geom_segment(aes(x = mean.prior, y = 0, xend = mean.prior, yend = dbeta(mean.prior, alpha, beta)),
                 size = 1, color = "#482173FF", linetype = "dashed") + # prior mean
    labs(x = "p",
         y = "Density",
         subtitle = "Density of the prior distribution and the prior mean in a dotted line") +
    xlim(c(0, 1)) +
    ylim(c(NA, 10)) +
    theme(text = element_text(size = 16),
          plot.subtitle = element_text(size = 14))
  
  
  
})

```








**Data and likelihood**
=====================================


Column {.sidebar data-width=350}
-------------------------------------



**Select dataset:**

```{r dataset}

selectInput("binom.data", label = "Select the data you want to analyze:",
            choices = c("Delta IV Heavy launches" = "delta",
                        "Dog cancer test detection" = "dog",
                        "Custom data" = "custom",
                        "Generate sample" = "generate"),
            selected = "delta")

```

Enter the value of $\hat{p}$ from your data

Enter the "true" value of $p$

```{r custom.phat}

numericInput("phat", label = withMathJax('\\( \\hat{p} \\text{ or } p \\):'),
            value = 0.5, min = 0, max = 1,
            step = 0.05)

```


Enter the value of the sample size $n$.



```{r custom.n}

numericInput("n", label = withMathJax('\\( n \\):'),
            value = 10, min = 0, max = 200,
            step = 1)

```


<hr>



The observed $\hat{p}$ is

```{r data.mean}
  
renderPrint({
  # read inputs
  alpha <- input$alpha
  beta <- input$beta
  dat.name <- input$binom.data
  phat <- input$phat
  n <- input$n
  
  # select data set
  if(dat.name == "dog"){
    binom.dat <- dog
    n <- nrow(binom.dat)
    y.lim <- 8
  } 
  
  if(dat.name == "delta"){
    binom.dat <- delta
    n <- nrow(binom.dat)
    y.lim <- 12
  } 
  
  if(dat.name == "custom"){
    binom.dat <- tibble(x = c(rep(1, round(n*phat, 0)), 
                              rep(0, n - round(n*phat, 0))))
    y.lim <- 10
  }
  
  if(dat.name == "generate"){
    binom.dat <- tibble(x = sample(c(0, 1), size = n, replace = TRUE, prob = c(1-phat, phat)))
    y.lim <- 10
  }
  
  print(mean(binom.dat$x))
  
  saveRDS(binom.dat, file = "binom_dat.RDS")
})  

```




Column
-------------------------------------


### 



```{r data.info}

renderPlot({

  # read inputs
  alpha <- input$alpha
  beta <- input$beta
  dat.name <- input$binom.data
  phat <- input$phat
  n <- input$n
  
  binom.dat <- readRDS("binom_dat.RDS")
  n <- nrow(binom.dat)
  y.lim <- 12
  
  
  colors <- c("Prior" = "#482173FF", 
              "Likelihood" = "#C2DF23FF", 
              "Posterior" = "#1E9B8AFF")
  
  
  # data
  data_plot <- binom.dat %>% 
    dplyr::mutate(x = factor(x, levels = c("1", "0"))) %>% 
    dplyr::group_by(x) %>% 
    dplyr::summarise(n = n()) %>% 
    dplyr::mutate(dat = "dat",
                  perc = n / nrow(binom.dat),
                  cumperc = cumsum(perc),
                  cum_n = cumsum(n),
                  cum_n = cum_n - n/2,
                  perc_label = paste0("x = ", x, ":\n", n, "\n", "(", round(100*perc, 1), "%)")) %>% 
    ggplot(aes(dat, n, fill = x)) +
    geom_bar(stat = "identity",
             position = "stack",
             fill="white",
             color = "black") +
    geom_text(aes(y = cum_n, label = perc_label),
              size = 6) +
    labs(x = "",
         y = "sample size n",
         subtitle = "Number and percentage of observed 0's and 1's") +
    theme(text = element_text(size = 16),
          plot.subtitle = element_text(size = 14),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.line.y = element_blank()) +
    coord_flip()
    
  
  
  
  
  # likelihood
  likelihood_plot <- ggplot(tibble(x = binom.dat %>% 
                               mutate(x = ifelse(x == 1, 0.97, 0.03)) %>% 
                               pull(), y = rep(0, n))) +
    stat_function(fun = dbeta, 
                  args = list(shape1 = sum(binom.dat$x) + 1, shape2 = n - sum(binom.dat$x) + 1),
                  aes(color = "Likelihood"),
                  size = 2,
                  n = 1000) + #adds likelihood density plot
    geom_segment(aes(x = sum(binom.dat$x) / n, y = 0, xend = sum(binom.dat$x) / n, yend = dbeta(sum(binom.dat$x) / n, sum(binom.dat$x) + 1, n - sum(binom.dat$x) + 1)),
                 color = "#C2DF23FF", size = 1, linetype = "dashed") + # data mean
    scale_color_manual(values = colors, breaks = c("Prior", "Likelihood", "Posterior")) +
    labs(x = "p",
         y = "",
         color = "",
         subtitle = expression(paste("Data likelihood in green and observed value of ", hat(p), " in a green dotted line"))) +
    xlim(c(0, 1)) +
    #ylim(c(NA, y.lim)) +
    theme(text = element_text(size = 16),
          plot.subtitle = element_text(size = 14))
  
    
   data_plot / likelihood_plot + 
  plot_layout(heights = c(1, 3))
    
  

})
```


<script>

function detect_change() {
    const paragraphs = document.getElementsByTagName('p');
    
    let index;
    for (var i=0; i<paragraphs.length; i++){
        if (paragraphs[i].innerHTML.includes('Enter the value of <span')){
            index = i;
            break;
        }
    }
    
    const n = document.querySelector("#n");
    
    const p_box = document.querySelector("#phat");
    p_label = document.querySelector("[for='phat']")
    switch (this.value) {
        case 'dog':
            n.value = 36;
            n.setAttribute("disabled", "disabled");
            p_box.style.display = 'None';
            p_label.style.display = 'None';
            paragraphs[index].style.display = 'None';
            paragraphs[index+1].style.display = 'None';
            paragraphs[index+2].style.display = 'None';
            break;
        case 'delta':
            n.value = 13;
            n.setAttribute("disabled", "disabled");
            p_box.style.display = 'None';
            p_label.style.display = 'None';
            paragraphs[index].style.display = 'None';
            paragraphs[index+1].style.display = 'None';
            paragraphs[index+2].style.display = 'None';
            break;
        case 'custom':
            n.removeAttribute("disabled");
            paragraphs[index].style.display = 'block';
            paragraphs[index+1].style.display = 'None';
            paragraphs[index+2].style.display = 'block';
            document.querySelector("#MathJax-Span-66").style.display = 'inline';
            document.querySelector("#MathJax-Span-67").style.display = 'None';
            document.querySelector("#MathJax-Span-68").style.display = 'None';
            p_box.style.display = 'block';
            p_label.style.display = 'block';
            p_box.value = 0.5;
            n.value = 10;
            break;
        case 'generate':
            n.removeAttribute("disabled");
            paragraphs[index].style.display = 'None';
            paragraphs[index+1].style.display = 'block';
            paragraphs[index+2].style.display = 'block';
            document.querySelector("#MathJax-Span-66").style.display = 'None';
            document.querySelector("#MathJax-Span-67").style.display = 'None';
            document.querySelector("#MathJax-Span-68").style.display = 'block';
            p_box.style.display = 'block';
            p_label.style.display = 'block';
            p_box.value = 0.5;
            n.value = 10;
    }
 }
 

$(document).ready(function() {
  //Add an event listener
  $(document).on("change", "select", detect_change)
  
  // trigger the event
  document.querySelector('select').dispatchEvent(new Event('change', { 'bubbles': true }))
});
</script>


**Posterior analysis**
=====================================


Column {.sidebar data-width=350}
-------------------------------------

**Posterior summary**

The table below contains the mean, quantiles, and variance of the posterior distribution.

```{r post.quantiles}




renderTable({
  # read inputs
  alpha <- input$alpha
  beta <- input$beta
  dat.name <- input$binom.data
  phat <- input$phat
  n <- input$n
  
  
  binom.dat <- readRDS("binom_dat.RDS")
  n <- nrow(binom.dat)
  y.lim <- 12
  
  
  # posterior values
  alpha.post <- alpha + sum(binom.dat$x)
  beta.post <- beta + n - sum(binom.dat$x)
  mean.post <- alpha.post / (alpha.post + beta.post)
  var.post <- alpha.post * beta.post / ((alpha.post + beta.post)^2 * (alpha.post + beta.post + 1))
  q1.post <- qbeta(p = 0.25, shape1 = alpha.post, shape2 = beta.post)
  median.post <- qbeta(p = 0.5, shape1 = alpha.post, shape2 = beta.post)
  q3.post <- qbeta(p = 0.75, shape1 = alpha.post, shape2 = beta.post)
  
  # summary table
  if(mean.post < median.post){
    summary <- tibble(" " = c("Q1", "Mean", "Median", "Q3", "Variance"),
                      Value = c(round(q1.post, 2), round(mean.post, 2), round(median.post, 2), round(q3.post, 2), as.character(round(var.post, 4))))
  }else{
    summary <- tibble(" " = c("Q1", "Median", "Mean", "Q3", "Variance"),
                      Value = c(round(q1.post, 2), round(median.post, 2), round(mean.post, 2), round(q3.post, 2), as.character(round(var.post, 4))))
  }
  
  print(summary)
  
})  
```


<hr>

**Credible intervals**

In the type-in box below, enter the level $\alpha$ of the credible interval:

```{r custom.alpha}

numericInput("significance", label = withMathJax('\\( \\alpha \\):'),
            value = 0.05, min = 0, max = 1,
            step = 0.01)

```



A $(1-\alpha) \times 100$% credible interval for $p$ is given by:



```{r post.ci}




renderTable({
  # read inputs
  alpha <- input$alpha
  beta <- input$beta
  dat.name <- input$binom.data
  phat <- input$phat
  n <- input$n
  significance <- input$significance
  
  binom.dat <- readRDS("binom_dat.RDS")
  n <- nrow(binom.dat)
  y.lim <- 12
  
  
  # posterior values
  alpha.post <- alpha + sum(binom.dat$x)
  beta.post <- beta + n - sum(binom.dat$x)
  
  # create tabe with CI
  ci <- betaHPD(alpha.post, beta.post, 1-significance)
  summary <- tibble(Lower = ci[1], Upper = ci[2])
  
  print(summary, digits = 4)
  
})  
```




Column
-------------------------------------

### 



```{r post.plot}

renderPlot({
  
  # read inputs
  alpha <- input$alpha
  beta <- input$beta
  dat.name <- input$binom.data
  phat <- input$phat
  n <- input$n
  significance <- input$significance
  
  binom.dat <- readRDS("binom_dat.RDS")
  n <- nrow(binom.dat)
  y.lim <- 12
  
  
  colors <- c("Prior" = "#482173FF", 
              "Likelihood" = "#C2DF23FF", 
              "Posterior" = "#1E9B8AFF")
  
  # posterior values
  alpha.post <- alpha + sum(binom.dat$x)
  beta.post <- beta + n - sum(binom.dat$x)
  
  # calculate posterior stats
  mean.post <- alpha.post / (alpha.post + beta.post)
  ci <- betaHPD(alpha.post, beta.post, 1-significance)
  
  ggplot(tibble(x = c(0, 1)), aes(x)) +
    stat_function(fun = dbeta, 
                  args = list(shape1 = alpha.post, shape2 = beta.post),
                  aes(color = "Posterior"),
                  size = 2,
                  n = 1000) + #adds posterior density plot
    geom_segment(aes(x = mean.post, y = 0, xend = mean.post, yend = dbeta(mean.post, alpha.post, beta.post)),
                 size = 1, color = "#1E9B8AFF", linetype = "dashed") + # posterior mean
    geom_segment(aes(x = ci[1], y = 0, xend = ci[1], yend = dbeta(ci[1], alpha.post, beta.post)),
                 linetype = "dotted", size = 1, color = "#1E9B8AFF") + # lower ci bound
    geom_segment(aes(x = ci[2], y = 0, xend = ci[2], yend = dbeta(ci[2], alpha.post, beta.post)),
                 linetype = "dotted", size = 1, color = "#1E9B8AFF") + # upper ci bound
    scale_color_manual(values = colors, breaks = c("Prior", "Likelihood", "Posterior")) +
    labs(x = "p",
         y = "Density",
         color = " ",
         subtitle = expression(paste("Density of the posterior distribution of p, posterior mean in a dashed line, ",  (1-alpha), " x100% credible interval in dashed lines."))) +
    xlim(c(0, 1)) +
    theme(text = element_text(size = 16),
          plot.subtitle = element_text(size = 14))
  
  
  
})
```








**Summary**
=====================================


Column {.sidebar data-width=350}
-------------------------------------


The table below contains summary statistics for the prior, data, and posterior.

```{r summary.table}

renderTable({
  # read inputs
  alpha <- input$alpha
  beta <- input$beta
  dat.name <- input$binom.data
  phat <- input$phat
  n <- input$n
  
  binom.dat <- readRDS("binom_dat.RDS")
  n <- nrow(binom.dat)
  y.lim <- 12
  
  # prior values
  mean.prior <- alpha / (alpha + beta)
  var.prior <- alpha * beta / ((alpha + beta)^2 * (alpha + beta + 1))
  
  # data values
  mean.data <- mean(binom.dat$x)
  var.data <- var(binom.dat$x)
  
  # posterior values
  alpha.post <- alpha + sum(binom.dat$x)
  beta.post <- beta + n - sum(binom.dat$x)
  mean.post <- alpha.post / (alpha.post + beta.post)
  var.post <- alpha.post * beta.post / ((alpha.post + beta.post)^2 * (alpha.post + beta.post + 1))
  
  # summary table
  summary <- tibble(" " = c("Prior", "Data", "Posterior"),
                    Mean = c(round(mean.prior, 2), round(mean.data, 2), round(mean.post, 2)),
                    Variance = c(round(var.prior, 4), "-", round(var.post, 4)))
  
  print(summary, digits = 2)
  
})  
```




<hr>

**Hyperparameter updates**

The table below contains the original and updated hyperparameters. The posterior distribution of $p$ is $\mathrm{Beta}(a', b')$, where $a'$ and $b'$ are the values in the second row.



```{r post.hyperparam.update}

renderTable({
  # read inputs
  alpha <- input$alpha
  beta <- input$beta
  dat.name <- input$binom.data
  phat <- input$phat
  n <- input$n
  significance <- input$significance
  
  binom.dat <- readRDS("binom_dat.RDS")
  n <- nrow(binom.dat)
  y.lim <- 12
  
  
  # posterior values
  alpha.post <- alpha + sum(binom.dat$x)
  beta.post <- beta + n - sum(binom.dat$x)
  
  # create tabe with CI
  summary <- tibble(" " = c("Original", "Updated"),
                    a = c(alpha, alpha.post),
                    b = c(beta, beta.post))
  
  print(summary, digits = 4)
  
})  
```








Column
-------------------------------------

### 



```{r summary.triplot}

renderPlot({
  
  # read inputs
  alpha <- input$alpha
  beta <- input$beta
  dat.name <- input$binom.data
  phat <- input$phat
  n <- input$n
  
  binom.dat <- readRDS("binom_dat.RDS")
  n <- nrow(binom.dat)
  y.lim <- 12
  
  
  colors <- c("Prior" = "#482173FF", 
              "Likelihood" = "#C2DF23FF", 
              "Posterior" = "#1E9B8AFF")
  
  
  # prior values
  mean.prior <- alpha / (alpha + beta)
  var.prior <- alpha * beta / ((alpha + beta)^2 * (alpha + beta + 1))
  
  # data values
  mean.data <- mean(binom.dat$x)
  var.data <- var(binom.dat$x)
  
  # posterior values
  alpha.post <- alpha + sum(binom.dat$x)
  beta.post <- beta + n - sum(binom.dat$x)
  mean.post <- alpha.post / (alpha.post + beta.post)
  var.post <- alpha.post * beta.post / ((alpha.post + beta.post)^2 * (alpha.post + beta.post + 1))
  
  
  
  ggplot(tibble(x = c(0, 1)), aes(x)) +
    stat_function(fun = dbeta, 
                  args = list(shape1 = alpha, shape2 = beta),
                  aes(color = "Prior"),
                  size = 2,
                  n = 1000) + #adds prior density plot
    geom_segment(aes(x = mean.prior, y = 0, xend = mean.prior, yend = dbeta(mean.prior, alpha, beta)),
                 size = 1, color = "#482173FF", linetype = "dashed") + # prior mean
    stat_function(fun = dbeta, 
                  args = list(shape1 = sum(binom.dat$x) + 1, shape2 = n - sum(binom.dat$x) + 1),
                  aes(color = "Likelihood"),
                  size = 2,
                  n = 1000) + #adds likelihood density plot
    geom_segment(aes(x = sum(binom.dat$x) / n, y = 0, xend = sum(binom.dat$x) / n, yend = dbeta(sum(binom.dat$x) / n, sum(binom.dat$x) + 1, n - sum(binom.dat$x) + 1)), 
                 color = "#C2DF23FF", size = 1, linetype = "dashed") + # data mean
    stat_function(fun = dbeta, 
                  args = list(shape1 = alpha + sum(binom.dat$x), shape2 = beta + n - sum(binom.dat$x)),
                  aes(color = "Posterior"),
                  size = 2,
                  n = 1000) + #adds posterior density plot
    geom_segment(aes(x = mean.post, y = 0, xend = mean.post, yend = dbeta(mean.post, alpha.post, beta.post)),
                 size = 1, color = "#1E9B8AFF", linetype = "dashed") + # posterior mean
    scale_color_manual(values = colors, breaks = c("Prior", "Likelihood", "Posterior")) +
    labs(x = "p",
         y = "Density",
         color = "Density:",
         subtitle = "Triplot: prior, likelihood, and posterior on the same display") +
    xlim(c(0, 1)) +
    theme(text = element_text(size = 16),
          plot.subtitle = element_text(size = 14))
  
  
  
})
```