---
title: "**First Bayes: Normal-Normal**"
author: ""
output: 
  flexdashboard::flex_dashboard:
    theme: lumen
    vertical_layout: fill
    orientation: columns
    source_code: https://github.com/GiankDiluvi/first-bayes-r
runtime: shiny
---

<style type="text/css">

.sidebar {  /* sidebar  */
   font-size: 14px;
   font-family: Helvetica;
   }
.navbar-inverse .navbar{
  font-family: Helvetica;
}
.navbar-inverse .navbar-brand{
  font-family: Helvetica;
  color: #000000;
  pointer-events: none;
}
</style>




```{r global, include=FALSE, warning = FALSE}
# preamble ###
# basics
library(flexdashboard)
library(tidyverse)
# plotting
library(ggplot2)
ggplot2::theme_set(theme_classic())
library(viridis)
library(patchwork)
# markdown
library(knitr)
library(kableExtra)
# shiny + html
library(shiny)
withMathJax()
library(DT)
# extras
#library(pscl)
set.seed(2020)
```


```{r}
normalHPD <- function(mu, sigma, p=0.95){
  # give a p% HPD credible interval for a N(mu, sigma) density
  
  out <- rep(0, 2)
  out[1] <- qnorm(p, mu, sigma, lower.tail = FALSE)
  out[2] <- qnorm(p, mu, sigma)
  out
}
```



```{r datasets}
# create data sets for analyses ###
iris <- tibble(x = datasets::iris$Sepal.Width) %>%  # sepal-width of iris data set
  head(22) # take first 22 rows
rock <- tibble(x = datasets::rock$area) %>% # area of rocks in petroleoum reservoir
  head(17) # take first 17 rows
```



**Prior distribution**
=====================================



Column {.sidebar data-width=350}
-------------------------------------

The prior distribution of $\mu$ is $\mathrm{\textsf{N}}(\mu_0, \sigma_0)$, and it characterizes your prior knowledge of $\mu$. The values $\mu_0$ and $\sigma_0$ are called the *hyperparameters*. 


**Select hyperparameters:**


```{r norm.hyperparams}

numericInput("mu0", label = withMathJax('Mean \\( \\mu_0 \\):'),
            min = -10000, max = 10000, value = 0)

numericInput("sigma0", label = withMathJax('Std. deviation \\( \\sigma_0 \\):'),
            min = 0, max = 10000, value = 1)


```

<hr>


With the values of $\mu_0$ and $\sigma_0$ that you chose, the prior expected value $\mathbb{E}[\mu]$ and variance $\mathrm{Var} [\mu]$ are shown below.


```{r prior.mean.var}

renderTable({
  
  # read inputs
  mu0 <- input$mu0
  sigma0 <- input$sigma0
  
  # calculate mean and variance
  mean <- mu0
  var <- sigma0^2
  
  # create table
  summary <- tibble(Mean = round(mean, 2),
                    Variance = round(var, 4))
  
  print(summary)
  
})

```





Column
-------------------------------------


### 
<!--
The plot below displays the density of the prior distribution with the hyperparameters that you chose.
-->

```{r prior}

renderPlot({
  
  # read inputs
  mu0 <- input$mu0
  sigma0 <- input$sigma0
  
  # calculate mean and variance
  tau <- 1 / sigma0^2
  xmin <- mu0 - 5*sigma0
  xmax <- mu0 + 5*sigma0
  ymax <- dnorm(mu0, mu0, sigma0)
  
  
  
  ggplot(tibble(x = c(xmin, xmax)), aes(x)) +
    stat_function(fun = dnorm, 
                  args = list(mean = mu0, sd = sigma0),
                  color = "#482173FF",
                  size = 2,
                  n = 1000) + #adds prior density plot
    geom_segment(aes(x = mu0, y = 0, xend = mu0, yend = ymax),
                 size = 1, color = "#482173FF", linetype = "dashed") + # prior mean
    labs(x = expression(mu),
         y = "Density",
         subtitle = "Density of the prior distribution and the prior mean in a dotted line") +
    xlim(c(xmin, xmax)) +
    ylim(c(NA, ymax)) +
    theme(text = element_text(size = 16),
          plot.subtitle = element_text(size = 14))
  
  
  
})

```








**Data and likelihood**
=====================================


Column {.sidebar data-width=350}
-------------------------------------



**Select dataset:**

```{r dataset}

selectInput("data", label = "Select the data you want to analyze:",
            choices = c("Iris sepal width" = "iris",
                        "Petroleoum reservoir rock areas" = "rock",
                        "Custom data" = "custom",
                        "Generate sample" = "generate"),
            selected = "iris")

```

Enter the value of $\hat{\mu}$ from your data

Enter the "true" value of $\mu$

```{r custom.muhat}

numericInput("muhat", label = "",
            value = 0, min = -10000, max = 10000)


```


Enter the value of the sample size $n$ and standard deviation $\sigma$.



```{r custom.n}

numericInput("n", label = withMathJax('\\( n \\):'),
            value = 10, min = 0, max = 10000,
            step = 1)
numericInput("sigma", label = withMathJax('\\( \\sigma \\):'),
            value = 1, min = 0, max = 10000)

```


<hr>



The observed $\hat{\mu}$ is

```{r data.mean}
  
renderPrint({
  # read inputs
  mu <- input$muhat
  sigma <- input$sigma
  dat.name <- input$data
  n <- input$n
  
  # select data set
  if(dat.name == "iris"){
    dat <- iris
    n <- nrow(dat)
  } 
  
  if(dat.name == "rock"){
    dat <- rock
    n <- nrow(dat)
  } 
  
  if(dat.name == "custom"){
    tmp_dat <- rnorm(n-1, mu, sigma)
    extra <- n*mu - sum(tmp_dat)
    dat <- tibble(x = c(tmp_dat, extra))
  }
  
  if(dat.name == "generate"){
    dat <- tibble(x = rnorm(n, mu, sigma))
  }
  
  print(mean(dat$x))
  
  saveRDS(dat, file = "norm_dat.RDS")
})  

```




Column
-------------------------------------


### 

<!--
The plot below displays the observed data as black points (when available), the data likelihood in green, and the observed value of $\hat{p}$ in a green dotted line.
-->


```{r data.info}

renderPlot({

  # read inputs
  mu <- input$muhat
  sigma <- input$sigma
  dat.name <- input$data
  n <- input$n
  
  # read data
  dat <- readRDS("norm_dat.RDS")
  n <- nrow(dat)
  
  # define sd and mean
  if(dat.name == "iris") sigma <- 0.4
  if(dat.name == "rock") sigma <- 1500
  xbar <- mean(dat$x)
  
  
  colors <- c("Prior" = "#482173FF", 
              "Likelihood" = "#C2DF23FF", 
              "Posterior" = "#1E9B8AFF")
  
  
  data_plot <- dat %>% 
    ggplot(aes(x)) +
    geom_histogram(color = "black", fill = "white") +
    xlim(c(xbar - 5*sigma, xbar + 5*sigma)) +
    labs(x = "Data",
         y = "Count",
         subtitle = "Histogram of the data") +
    theme(text = element_text(size = 16),
          plot.subtitle = element_text(size = 14))
  
  
  # likelihood
  likelihood_plot <- dat %>% 
    ggplot() +
    stat_function(fun = dnorm, 
                  args = list(mean = mean(dat$x), sd = sigma),
                  aes(color = "Likelihood"),
                  size = 2,
                  n = 1000) + #adds likelihood density plot
    geom_segment(aes(x = xbar, y = 0, xend = xbar, yend = dnorm(xbar, xbar, sigma)),
                 color = "#C2DF23FF", size = 1, linetype = "dashed") + # data mean
    scale_color_manual(values = colors, breaks = c("Prior", "Likelihood", "Posterior")) +
    labs(x = expression(mu),
         y = "",
         color = "",
         subtitle = expression(paste("Data likelihood in green and observed value of ", hat(mu), " in a green dotted line"))) +
    xlim(c(xbar - 5*sigma, xbar + 5*sigma)) +
    ylim(c(NA, dnorm(xbar, xbar, sigma)*1.1)) +
    theme(text = element_text(size = 16),
          plot.subtitle = element_text(size = 14))
  
  
    
   data_plot / likelihood_plot + 
  plot_layout(heights = c(2, 3))
    
  

})
```


<script>

function detect_change() {
    const paragraphs = document.getElementsByTagName('p');
    
    let index;
    for (var i=0; i<paragraphs.length; i++){
        if (paragraphs[i].innerHTML.includes('Enter the value of <span')){
            index = i;
            break;
        }
    }
    
    const n = document.querySelector("#n");
    const sigma = document.querySelector("#sigma");
    
    const mu_box = document.querySelector("#muhat");
    mu_label = document.querySelector("[for='muhat']")
    switch (this.value) {
        case 'iris':
            n.value = 22;
            n.setAttribute("disabled", "disabled");
            sigma.value = 0.4;
            sigma.setAttribute("disabled", "disabled");
            mu_box.style.display = 'None';
            mu_label.style.display = 'None';
            paragraphs[index].style.display = 'None';
            paragraphs[index+1].style.display = 'None';
            paragraphs[index+2].style.display = 'None';
            break;
        case 'rock':
            n.value = 17;
            n.setAttribute("disabled", "disabled");
            sigma.value = 1500;
            sigma.setAttribute("disabled", "disabled");
            p_box.style.display = 'None';
            p_label.style.display = 'None';
            paragraphs[index].style.display = 'None';
            paragraphs[index+1].style.display = 'None';
            paragraphs[index+2].style.display = 'None';
            break;
        case 'custom':
            n.removeAttribute("disabled");
            sigma.removeAttribute("disabled");
            paragraphs[index].style.display = 'block';
            paragraphs[index+1].style.display = 'None';
            paragraphs[index+2].style.display = 'block';
            document.querySelector("#MathJax-Span-66").style.display = 'inline';
            document.querySelector("#MathJax-Span-67").style.display = 'None';
            document.querySelector("#MathJax-Span-68").style.display = 'None';
            mu_box.style.display = 'block';
            mu_label.style.display = 'block';
            mu_box.value = 0.5;
            n.value = 10;
            sigma.value = 1;
            break;
        case 'generate':
            n.removeAttribute("disabled");
            sigma.removeAttribute("disabled");
            paragraphs[index].style.display = 'None';
            paragraphs[index+1].style.display = 'block';
            paragraphs[index+2].style.display = 'block';
            document.querySelector("#MathJax-Span-66").style.display = 'None';
            document.querySelector("#MathJax-Span-67").style.display = 'None';
            document.querySelector("#MathJax-Span-68").style.display = 'block';
            mu_box.style.display = 'block';
            mu_label.style.display = 'block';
            mu_box.value = 0.5;
            n.value = 10;
            sigma.value = 1;
    }
 }
 

$(document).ready(function() {
  //Add an event listener
  $(document).on("change", "select", detect_change)
  
  // trigger the event
  document.querySelector('select').dispatchEvent(new Event('change', { 'bubbles': true }))
});
</script>


**Posterior analysis**
=====================================


Column {.sidebar data-width=350}
-------------------------------------

**Posterior summary**

The table below contains the mean, quantiles, and variance of the posterior distribution.

```{r post.quantiles}




renderTable({
  # read inputs
  mu0 <- input$mu0
  sigma0 <- input$sigma0
  mu <- input$muhat
  sigma <- input$sigma
  dat.name <- input$data
  n <- input$n
  
  
  # read data
  dat <- readRDS("norm_dat.RDS")
  n <- nrow(dat)
  
  if(dat.name == "iris") sigma <- 0.4
  if(dat.name == "rock") sigma <- 1500
  
  
  # posterior values
  xbar <- mean(dat$x)
  tau0 <- 1 / sigma0^2
  tau <- 1 / sigma^2
  tau1 <- tau0 + n*tau
  mu1 <- (tau0*mu0 + n*tau*xbar) / tau1
  sigma1 <- 1 / sqrt(tau1)
  
  # define table values
  mean.post <- mu1
  var.post <- sigma1^2
  q1.post <- qnorm(p = 0.25, mean = mu1, sd = sigma1)
  median.post <- qnorm(p = 0.5, mean = mu1, sd = sigma1)
  q3.post <- qnorm(p = 0.75, mean = mu1, sd = sigma1)
  
  # summary table
  if(mean.post < median.post){
    summary <- tibble(" " = c("Q1", "Mean", "Median", "Q3", "Variance"),
                      Value = c(round(q1.post, 2), round(mean.post, 2), round(median.post, 2), round(q3.post, 2), as.character(round(var.post, 4))))
  }else{
    summary <- tibble(" " = c("Q1", "Median", "Mean", "Q3", "Variance"),
                      Value = c(round(q1.post, 2), round(median.post, 2), round(mean.post, 2), round(q3.post, 2), as.character(round(var.post, 4))))
  }
  
  print(summary)
  
})  
```


<hr>

**Credible intervals**

In the type-in box below, enter the level $\alpha$ of the credible interval:

```{r custom.alpha}

numericInput("significance", label = withMathJax('\\( \\alpha \\):'),
            value = 0.05, min = 0, max = 1,
            step = 0.01)

```



A $(1-\alpha) \times 100$% credible interval for $\mu$ is given by:



```{r post.ci}




renderTable({
  # read inputs
  mu0 <- input$mu0
  sigma0 <- input$sigma0
  mu <- input$muhat
  sigma <- input$sigma
  dat.name <- input$data
  n <- input$n
  significance <- input$significance
  
  
  # read data
  dat <- readRDS("norm_dat.RDS")
  n <- nrow(dat)
  
  if(dat.name == "iris") sigma <- 0.4
  if(dat.name == "rock") sigma <- 1500
  
  
  # posterior values
  xbar <- mean(dat$x)
  tau0 <- 1 / sigma0^2
  tau <- 1 / sigma^2
  tau1 <- tau0 + n*tau
  mu1 <- (tau0*mu0 + n*tau*xbar) / tau1
  sigma1 <- 1 / sqrt(tau1)
  
  # create tabe with CI
  ci <- normalHPD(mu1, sigma1, 1-significance)
  summary <- tibble(Lower = ci[1], Upper = ci[2])
  
  print(summary, digits = 4)
  
})  
```




Column
-------------------------------------

### 

<!--
Density of the posterior distribution of $p$, posterior mean in a dashed line, $(1-\alpha) \times 100$% credible interval in dashed lines.
-->


```{r post.plot}

renderPlot({
  
  # read inputs
  mu0 <- input$mu0
  sigma0 <- input$sigma0
  mu <- input$muhat
  sigma <- input$sigma
  dat.name <- input$data
  n <- input$n
  significance <- input$significance
  
  
  # read data
  dat <- readRDS("norm_dat.RDS")
  n <- nrow(dat)
  
  if(dat.name == "iris") sigma <- 0.4
  if(dat.name == "rock") sigma <- 1500
  
  
  # posterior values
  xbar <- mean(dat$x)
  tau0 <- 1 / sigma0^2
  tau <- 1 / sigma^2
  tau1 <- tau0 + n*tau
  mu1 <- (tau0*mu0 + n*tau*xbar) / tau1
  sigma1 <- 1 / sqrt(tau1)
  ci <- normalHPD(mu1, sigma1, 1-significance)
  
  
  colors <- c("Prior" = "#482173FF", 
              "Likelihood" = "#C2DF23FF", 
              "Posterior" = "#1E9B8AFF")
  

  
  
  ggplot(tibble(x = c(mu1 - 5*sigma1, mu1 + 5*sigma1)), aes(x)) +
    stat_function(fun = dnorm, 
                  args = list(mean = mu1, sd = sigma1),
                  aes(color = "Posterior"),
                  size = 2,
                  n = 1000) + #adds posterior density plot
    geom_segment(aes(x = mu1, y = 0, xend = mu1, yend = dnorm(mu1, mu1, sigma1)),
                 size = 1, color = "#1E9B8AFF", linetype = "dashed") + # posterior mean
    geom_segment(aes(x = ci[1], y = 0, xend = ci[1], yend = dnorm(ci[1], mu1, sigma1)),
                 linetype = "dotted", size = 1, color = "#1E9B8AFF") + # lower ci bound
    geom_segment(aes(x = ci[2], y = 0, xend = ci[2], yend = dnorm(ci[2], mu1, sigma1)),
                 linetype = "dotted", size = 1, color = "#1E9B8AFF") + # upper ci bound
    scale_color_manual(values = colors, breaks = c("Prior", "Likelihood", "Posterior")) +
    labs(x = expression(mu),
         y = "Density",
         color = " ",
         subtitle = expression(paste("Density of the posterior distribution of ", mu, ", posterior mean in a dashed line, ",  (1-alpha), " x100% credible interval in dashed lines."))) +
    xlim(c(mu1 - 5*sigma1, mu1 + 5*sigma1)) +
    theme(text = element_text(size = 16),
          plot.subtitle = element_text(size = 14))
  
  
  
})
```








**Summary**
=====================================


Column {.sidebar data-width=350}
-------------------------------------


The table below contains summary statistics for the prior, data, and posterior.

```{r summary.table}

renderTable({
  # read inputs
  mu0 <- input$mu0
  sigma0 <- input$sigma0
  mu <- input$muhat
  sigma <- input$sigma
  dat.name <- input$data
  n <- input$n
  
  
  # read data
  dat <- readRDS("norm_dat.RDS")
  n <- nrow(dat)
  
  if(dat.name == "iris") sigma <- 0.4
  if(dat.name == "rock") sigma <- 1500
  
  
  # posterior values
  xbar <- mean(dat$x)
  tau0 <- 1 / sigma0^2
  tau <- 1 / sigma^2
  tau1 <- tau0 + n*tau
  mu1 <- (tau0*mu0 + n*tau*xbar) / tau1
  sigma1 <- 1 / sqrt(tau1)
  
  # prior values
  mean.prior <- mu0
  var.prior <- sigma0^2
  
  # data values
  mean.data <- mean(dat$x)
  var.data <- var(dat$x)
  
  # posterior values
  mean.post <- mu1
  var.post <- sigma1^2
  
  # summary table
  summary <- tibble(" " = c("Prior", "Data", "Posterior"),
                    Mean = c(round(mean.prior, 2), round(mean.data, 2), round(mean.post, 2)),
                    Variance = c(round(var.prior, 4), "-", round(var.post, 4)))
  
  print(summary, digits = 2)
  
})  
```




<hr>

**Hyperparameter updates**

The table below contains the original and updated hyperparameters. The posterior distribution of $\mu$ is $\mathrm{\textsf{N}}(\mu_1, \sigma_1^2)$, where $\mu_1$ and $\sigma_1$ are the values in the second row.



```{r post.hyperparam.update}

renderTable({
  # read inputs
  mu0 <- input$mu0
  sigma0 <- input$sigma0
  mu <- input$muhat
  sigma <- input$sigma
  dat.name <- input$data
  n <- input$n
  
  
  # read data
  dat <- readRDS("norm_dat.RDS")
  n <- nrow(dat)
  
  if(dat.name == "iris") sigma <- 0.4
  if(dat.name == "rock") sigma <- 1500
  
  
  # posterior values
  xbar <- mean(dat$x)
  tau0 <- 1 / sigma0^2
  tau <- 1 / sigma^2
  tau1 <- tau0 + n*tau
  mu1 <- (tau0*mu0 + n*tau*xbar) / tau1
  sigma1 <- 1 / sqrt(tau1)
  
  
  # create tabe with CI
  summary <- tibble(" " = c("Original", "Updated"),
                    mean = c(mu0, mu1),
                    sd = c(sigma0, sigma1))
  
  print(summary, digits = 4)
  
})  
```








Column
-------------------------------------

### 



```{r summary.triplot}

renderPlot({
  
  # read inputs
  mu0 <- input$mu0
  sigma0 <- input$sigma0
  mu <- input$muhat
  sigma <- input$sigma
  dat.name <- input$data
  n <- input$n
  
  
  # read data
  dat <- readRDS("norm_dat.RDS")
  n <- nrow(dat)
  
  if(dat.name == "iris") sigma <- 0.4
  if(dat.name == "rock") sigma <- 1500
  
  
  # posterior values
  xbar <- mean(dat$x)
  tau0 <- 1 / sigma0^2
  tau <- 1 / sigma^2
  tau1 <- tau0 + n*tau
  mu1 <- (tau0*mu0 + n*tau*xbar) / tau1
  sigma1 <- 1 / sqrt(tau1)
  
  # prior values
  mean.prior <- mu0
  var.prior <- sigma0^2
  
  # data values
  mean.data <- mean(dat$x)
  var.data <- var(dat$x)
  
  # posterior values
  mean.post <- mu1
  var.post <- sigma1^2
  
  # limits
  if(sigma0 > 100*sigma1){
    # if prior variance is super large, then prior is considered non-informative
    xmin <- min(xbar - 3*sigma, mu1 - 3*sigma1)
    xmax <- max(xbar + 3*sigma, mu1 + 3*sigma1)
  }else{
    # if prior is informative, then plot correctly
    xmin <- min(mu0 - 3*sigma0, xbar - 3*sigma, mu1 - 3*sigma1)
    xmax <- max(mu0 + 3*sigma0, xbar + 3*sigma, mu1 + 3*sigma1)
  }
  
  
  
  
  colors <- c("Prior" = "#482173FF", 
              "Likelihood" = "#C2DF23FF", 
              "Posterior" = "#1E9B8AFF")
  
  
  ggplot(tibble(x = c(xmin, xmax)), aes(x)) +
    stat_function(fun = dnorm, 
                  args = list(mean = mu0, sd = sigma0),
                  aes(color = "Prior"),
                  size = 2,
                  n = 1000) + #adds prior density plot
    geom_segment(aes(x = mean.prior, y = 0, xend = mean.prior, yend = dnorm(mean.prior, mu0, sigma0)),
                 size = 1, color = "#482173FF", linetype = "dashed") + # prior mean
    stat_function(fun = dnorm, 
                  args = list(mean = xbar, sd = sigma),
                  aes(color = "Likelihood"),
                  size = 2,
                  n = 1000) + #adds likelihood density plot
    geom_segment(aes(x = xbar, y = 0, xend = xbar, yend = dnorm(xbar, xbar, sigma)), 
                 color = "#C2DF23FF", size = 1, linetype = "dashed") + # data mean
    stat_function(fun = dnorm, 
                  args = list(mean = mu1, sd = sigma1),
                  aes(color = "Posterior"),
                  size = 2,
                  n = 1000) + #adds posterior density plot
    geom_segment(aes(x = mean.post, y = 0, xend = mean.post, yend = dnorm(mean.post, mu1, sigma1)),
                 size = 1, color = "#1E9B8AFF", linetype = "dashed") + # posterior mean
    scale_color_manual(values = colors, breaks = c("Prior", "Likelihood", "Posterior")) +
    labs(x = expression(mu),
         y = "Density",
         color = "Density:",
         subtitle = "Triplot: prior, likelihood, and posterior on the same display") +
    xlim(c(xmin, xmax)) +
    theme(text = element_text(size = 16),
          plot.subtitle = element_text(size = 14))
  
  
  
})
```




